---
title: "ESPN basketball endpoints -- WBB & WNBA"
author: "Saiem Gilani"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ESPN basketball endpoints}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = "#>", collapse = TRUE)
```

## Introduction

wehoop wraps three ESPN API families for women's basketball -- no authentication
required for any of them:

- **site-v2** (`site.api.espn.com/apis/site/v2/sports/basketball/{league}`) --
  scoreboards, teams, rosters, schedules, news, injuries, standings, player and
  team stats. The most stable surface.
- **core-v2** (`sports.core.api.espn.com/v2/sports/basketball/{league}`) --
  athletes index, event detail (odds, probabilities, officials, broadcasts),
  seasons catalog, venues, coaches, draft/transactions (WNBA). Uses `$ref`
  pagination for large collections.
- **web-common-v3** (`site.web.api.espn.com/apis/common/v3/sports/basketball/{league}`) --
  per-athlete deep detail: overview, stats, gamelog, splits. Less stable than
  site-v2; some legacy seasons return 404.

A fourth host, `cdn.espn.com/core`, exists but is fully redundant with
site-v2 `/summary` parsing and is not wrapped by wehoop.

League slugs used internally:

- WBB: `basketball/womens-college-basketball`
- WNBA: `basketball/wnba`

**Rate limiting.** ESPN does not publish rate limits, but requests that arrive
faster than one per second occasionally receive HTTP 429 or silent empty
responses. Add `Sys.sleep(1)` between calls in scripts that loop over many
games or athletes.

**Proxy.** ESPN wrappers call `.retry_request()` directly. Per-call proxy
override via `...` is not supported for these functions. Use
`options(wehoop.proxy = "http://host:port")` (or a named list for
authenticated proxies) once at the top of your session and every ESPN call
will route through it automatically.

## Endpoint surface map

### Scoreboard, teams, play-by-play, and box score (original 22)

| Function | Returns | Example call |
|---|---|---|
| `espn_wbb_scoreboard()` | Daily scoreboard with game status | `espn_wbb_scoreboard(season = 2025)` |
| `espn_wbb_teams()` | All WBB teams | `espn_wbb_teams()` |
| `espn_wbb_conferences()` | Conference list | `espn_wbb_conferences()` |
| `espn_wbb_rankings()` | AP/coaches poll rankings | `espn_wbb_rankings()` |
| `espn_wbb_standings()` | Conference standings | `espn_wbb_standings(season = 2025)` |
| `espn_wbb_team_stats()` | Team-level season stats | `espn_wbb_team_stats(season = 2025)` |
| `espn_wbb_player_stats()` | Player-level season stats | `espn_wbb_player_stats(season = 2025)` |
| `espn_wbb_game_all()` | Full game summary (named list) | `espn_wbb_game_all(game_id = "401768729")` |
| `espn_wbb_pbp()` | Play-by-play | `espn_wbb_pbp(game_id = "401768729")` |
| `espn_wbb_team_box()` | Team box score | `espn_wbb_team_box(game_id = "401768729")` |
| `espn_wbb_player_box()` | Player box score | `espn_wbb_player_box(game_id = "401768729")` |
| `espn_wbb_game_rosters()` | Game-day rosters | `espn_wbb_game_rosters(game_id = "401768729")` |
| `espn_wnba_scoreboard()` | Daily WNBA scoreboard | `espn_wnba_scoreboard(season = 2025)` |
| `espn_wnba_teams()` | All WNBA teams | `espn_wnba_teams()` |
| `espn_wnba_standings()` | WNBA standings | `espn_wnba_standings(season = 2025)` |
| `espn_wnba_team_stats()` | Team-level season stats | `espn_wnba_team_stats(season = 2025)` |
| `espn_wnba_player_stats()` | Player-level season stats | `espn_wnba_player_stats(season = 2025)` |
| `espn_wnba_game_all()` | Full game summary (named list) | `espn_wnba_game_all(game_id = "401736171")` |
| `espn_wnba_pbp()` | Play-by-play | `espn_wnba_pbp(game_id = "401736171")` |
| `espn_wnba_team_box()` | Team box score | `espn_wnba_team_box(game_id = "401736171")` |
| `espn_wnba_player_box()` | Player box score | `espn_wnba_player_box(game_id = "401736171")` |
| `espn_wnba_game_rosters()` | Game-day rosters | `espn_wnba_game_rosters(game_id = "401736171")` |

### Phase 1 -- Reference & catalog (18 new)

| Function | Returns | Example call |
|---|---|---|
| `espn_wbb_news()` | League news feed | `espn_wbb_news(limit = 50)` |
| `espn_wbb_calendar()` | Season calendar weeks | `espn_wbb_calendar(season = 2025)` |
| `espn_wbb_team()` | Single-team detail (info, record, coaches) | `espn_wbb_team(team_id = 2509, season = 2025)` |
| `espn_wbb_team_roster()` | Team roster | `espn_wbb_team_roster(team_id = 2509, season = 2025)` |
| `espn_wbb_team_schedule()` | Team schedule | `espn_wbb_team_schedule(team_id = 2509, season = 2025)` |
| `espn_wbb_team_news()` | Team news feed | `espn_wbb_team_news(team_id = 2509, limit = 25)` |
| `espn_wbb_team_leaders()` | Team statistical leaders | `espn_wbb_team_leaders(team_id = 2509, season = 2025)` |
| `espn_wbb_injuries()` | League-wide injury report | `espn_wbb_injuries(season = 2025)` |
| `espn_wbb_team_injuries()` | Team injury report | `espn_wbb_team_injuries(team_id = 2509)` |
| `espn_wnba_news()` | League news feed | `espn_wnba_news(limit = 50)` |
| `espn_wnba_calendar()` | Season calendar weeks | `espn_wnba_calendar(season = 2025)` |
| `espn_wnba_team()` | Single-team detail (info, record, coaches) | `espn_wnba_team(team_id = 17, season = 2025)` |
| `espn_wnba_team_roster()` | Team roster | `espn_wnba_team_roster(team_id = 17, season = 2025)` |
| `espn_wnba_team_schedule()` | Team schedule | `espn_wnba_team_schedule(team_id = 17, season = 2025)` |
| `espn_wnba_team_news()` | Team news feed | `espn_wnba_team_news(team_id = 17, limit = 25)` |
| `espn_wnba_team_leaders()` | Team statistical leaders | `espn_wnba_team_leaders(team_id = 17, season = 2025)` |
| `espn_wnba_injuries()` | League-wide injury report | `espn_wnba_injuries(season = 2025)` |
| `espn_wnba_team_injuries()` | Team injury report | `espn_wnba_team_injuries(team_id = 17)` |

### Phase 2 -- Athlete coverage (16 new)

| Function | Returns | Example call |
|---|---|---|
| `espn_wbb_athlete_info()` | Athlete bio and profile | `espn_wbb_athlete_info(athlete_id = "4433403")` |
| `espn_wbb_athlete_overview()` | Season overview (web-common-v3) | `espn_wbb_athlete_overview(athlete_id = "4433403", season = 2025)` |
| `espn_wbb_athlete_stats()` | Season stats by category | `espn_wbb_athlete_stats(athlete_id = "4433403", season = 2025)` |
| `espn_wbb_athlete_gamelog()` | Game-by-game log | `espn_wbb_athlete_gamelog(athlete_id = "4433403", season = 2025)` |
| `espn_wbb_athlete_splits()` | Splits by situation | `espn_wbb_athlete_splits(athlete_id = "4433403", season = 2025)` |
| `espn_wbb_athlete_eventlog()` | Event log ($ref links) | `espn_wbb_athlete_eventlog(athlete_id = "4433403", season = 2025)` |
| `espn_wbb_athlete_awards()` | Awards history | `espn_wbb_athlete_awards(athlete_id = "4433403")` |
| `espn_wbb_athlete_statisticslog()` | Statistics log (core-v2) | `espn_wbb_athlete_statisticslog(athlete_id = "4433403", season = 2025)` |
| `espn_wnba_athlete_info()` | Athlete bio and profile | `espn_wnba_athlete_info(athlete_id = "3149391")` |
| `espn_wnba_athlete_overview()` | Season overview (web-common-v3) | `espn_wnba_athlete_overview(athlete_id = "3149391", season = 2025)` |
| `espn_wnba_athlete_stats()` | Season stats by category | `espn_wnba_athlete_stats(athlete_id = "3149391", season = 2025)` |
| `espn_wnba_athlete_gamelog()` | Game-by-game log | `espn_wnba_athlete_gamelog(athlete_id = "3149391", season = 2025)` |
| `espn_wnba_athlete_splits()` | Splits by situation | `espn_wnba_athlete_splits(athlete_id = "3149391", season = 2025)` |
| `espn_wnba_athlete_eventlog()` | Event log ($ref links) | `espn_wnba_athlete_eventlog(athlete_id = "3149391", season = 2025)` |
| `espn_wnba_athlete_awards()` | Awards history | `espn_wnba_athlete_awards(athlete_id = "3149391")` |
| `espn_wnba_athlete_statisticslog()` | Statistics log (core-v2) | `espn_wnba_athlete_statisticslog(athlete_id = "3149391", season = 2025)` |

### Phase 3 -- Event enrichment (8 new)

| Function | Returns | Example call |
|---|---|---|
| `espn_wbb_event_odds()` | Opening/closing lines | `espn_wbb_event_odds(event_id = "401768729")` |
| `espn_wbb_event_probabilities()` | Win probability by play | `espn_wbb_event_probabilities(event_id = "401768729")` |
| `espn_wbb_event_officials()` | Game officials | `espn_wbb_event_officials(event_id = "401768729")` |
| `espn_wbb_event_broadcasts()` | TV/radio broadcast info | `espn_wbb_event_broadcasts(event_id = "401768729")` |
| `espn_wnba_event_odds()` | Opening/closing lines | `espn_wnba_event_odds(event_id = "401736171")` |
| `espn_wnba_event_probabilities()` | Win probability by play | `espn_wnba_event_probabilities(event_id = "401736171")` |
| `espn_wnba_event_officials()` | Game officials | `espn_wnba_event_officials(event_id = "401736171")` |
| `espn_wnba_event_broadcasts()` | TV/radio broadcast info | `espn_wnba_event_broadcasts(event_id = "401736171")` |

### Phase 4 -- WNBA-only artifacts (3 new)

| Function | Returns | Example call |
|---|---|---|
| `espn_wnba_draft()` | Draft picks by season | `espn_wnba_draft(season = 2025)` |
| `espn_wnba_freeagents()` | Free agent list | `espn_wnba_freeagents(season = 2025)` |
| `espn_wnba_transactions()` | Transactions log | `espn_wnba_transactions(season = 2025, limit = 100)` |

### Phase 5 -- League-wide catalogs (12 new)

| Function | Returns | Example call |
|---|---|---|
| `espn_wbb_leaders()` | Statistical leaders | `espn_wbb_leaders(season = 2025)` |
| `espn_wbb_venues()` | All arenas | `espn_wbb_venues()` |
| `espn_wbb_coaches()` | All coaches | `espn_wbb_coaches(season = 2025)` |
| `espn_wbb_athletes_index()` | Full athlete index | `espn_wbb_athletes_index(season = 2025, limit = 50)` |
| `espn_wbb_seasons()` | All seasons on record | `espn_wbb_seasons()` |
| `espn_wbb_season_info()` | Single season metadata | `espn_wbb_season_info(season = 2025)` |
| `espn_wnba_leaders()` | Statistical leaders | `espn_wnba_leaders(season = 2025)` |
| `espn_wnba_venues()` | All arenas | `espn_wnba_venues()` |
| `espn_wnba_coaches()` | All coaches | `espn_wnba_coaches(season = 2025)` |
| `espn_wnba_athletes_index()` | Full athlete index | `espn_wnba_athletes_index(season = 2025, limit = 50)` |
| `espn_wnba_seasons()` | All seasons on record | `espn_wnba_seasons()` |
| `espn_wnba_season_info()` | Single season metadata | `espn_wnba_season_info(season = 2025)` |

### Phase 6 -- WNBA symmetry (1 new)

| Function | Returns | Example call |
|---|---|---|
| `espn_wnba_conferences()` | WNBA conference list | `espn_wnba_conferences()` |

## Phase 1 walkthrough -- reference and catalog

These examples use UConn (`team_id = 2509`) for WBB and the Las Vegas Aces
(`team_id = 17`) for WNBA.

### League news

```{r phase1-news}
library(wehoop)

# Latest 10 WBB news items
wbb_news <- espn_wbb_news(limit = 10)
head(wbb_news[, c("headline", "published")])

# Latest 10 WNBA news items
wnba_news <- espn_wnba_news(limit = 10)
head(wnba_news[, c("headline", "published")])
```

### Season calendar

```{r phase1-calendar}
# WBB 2025 season calendar
wbb_cal <- espn_wbb_calendar(season = 2025)
head(wbb_cal)

# WNBA 2025 season calendar
wnba_cal <- espn_wnba_calendar(season = 2025)
head(wnba_cal)
```

### Team detail, roster, and schedule

```{r phase1-team}
# UConn team info (named list: Info, Record, Coaches, ...)
uconn <- espn_wbb_team(team_id = 2509, season = 2025)
names(uconn)

# UConn 2025 roster
uconn_roster <- espn_wbb_team_roster(team_id = 2509, season = 2025)
head(uconn_roster[, c("athlete_display_name", "position_name", "jersey")])

# UConn 2025 schedule (regular season)
uconn_sched <- espn_wbb_team_schedule(team_id = 2509, season = 2025, season_type = 2)
head(uconn_sched[, c("name", "date", "home_away", "score")])

# Las Vegas Aces roster
aces_roster <- espn_wnba_team_roster(team_id = 17, season = 2025)
head(aces_roster[, c("athlete_display_name", "position_name", "jersey")])
```

### Team news and statistical leaders

```{r phase1-leaders}
# UConn team news
uconn_news <- espn_wbb_team_news(team_id = 2509, limit = 5)
head(uconn_news[, c("headline", "published")])

# UConn statistical leaders for 2025
uconn_ldrs <- espn_wbb_team_leaders(team_id = 2509, season = 2025)
head(uconn_ldrs)

# Las Vegas Aces leaders
aces_ldrs <- espn_wnba_team_leaders(team_id = 17, season = 2025)
head(aces_ldrs)
```

### Injury reports

```{r phase1-injuries}
# WNBA league-wide injuries
wnba_inj <- espn_wnba_injuries(season = 2025)
# WBB injuries are sparse on ESPN; an empty tibble is normal
wbb_inj <- espn_wbb_injuries(season = 2025)

# Team-level injury report
aces_inj <- espn_wnba_team_injuries(team_id = 17)
```

## Phase 2 walkthrough -- athlete deep dive

The pattern: pull a roster, pick an athlete ID, then call the athlete-
specific endpoints in sequence.

```{r phase2-roster-to-id}
library(wehoop)
library(dplyr)

# Step 1 -- get UConn roster
uconn_roster <- espn_wbb_team_roster(team_id = 2509, season = 2025)

# Step 2 -- pick a player (first one returned)
athlete_id <- uconn_roster$athlete_id[1]
athlete_name <- uconn_roster$athlete_display_name[1]
message("Selected: ", athlete_name, " (", athlete_id, ")")
```

```{r phase2-info}
# Step 3 -- bio and profile
bio <- espn_wbb_athlete_info(athlete_id = athlete_id)
glimpse(bio)
```

```{r phase2-overview}
# Step 4 -- season overview (web-common-v3)
overview <- espn_wbb_athlete_overview(athlete_id = athlete_id, season = 2025)
# overview is a named list; main tables vary by season/athlete
names(overview)
```

```{r phase2-gamelog}
# Step 5 -- game-by-game log
gamelog <- espn_wbb_athlete_gamelog(athlete_id = athlete_id, season = 2025)
head(gamelog[, c("game_date", "opponent", "points", "rebounds", "assists")])
```

```{r phase2-splits}
# Step 6 -- splits (home/away, wins/losses, by month, ...)
splits <- espn_wbb_athlete_splits(athlete_id = athlete_id, season = 2025)
head(splits)
```

The same chain works for WNBA; swap `espn_wbb_*` for `espn_wnba_*` and use
a WNBA `athlete_id` (e.g., A'ja Wilson is `"3149391"`).

### Notes on Phase 2 edge cases

- `espn_*_athlete_eventlog()` returns `$ref` URL columns pointing to
  individual game stat resources. These are not auto-resolved; use
  `espn_*_athlete_gamelog()` for pre-parsed per-game stats.
- `espn_*_athlete_awards()` is frequently empty. Many athletes have no
  ESPN-recorded awards; an empty tibble is the normal return, not an error.
- `espn_*_athlete_statisticslog()` (core-v2) and `_overview()` (web-common-v3)
  may return HTTP 404 for seasons prior to roughly 2018. Use
  `espn_*_athlete_gamelog()` for historical seasons.

## Phase 3 walkthrough -- event enrichment

### Win-probability chart (WNBA)

Combine play-by-play with win probabilities to visualize game momentum.
Event `'401736171'` is a 2024 WNBA regular-season game.

```{r phase3-prob}
library(wehoop)
library(dplyr)

game_id <- "401736171"

# Play-by-play (includes clock and score for labelling)
pbp <- espn_wnba_pbp(game_id = game_id)

# Win probabilities from core-v2 (one row per scored play, up to limit)
probs <- espn_wnba_event_probabilities(event_id = game_id, limit = 200)
head(probs[, c("sequence_number", "home_win_percentage", "away_win_percentage")])
```

```{r phase3-plot}
library(ggplot2)

# Merge by sequence number and plot
plot_data <- probs %>%
  mutate(seq = as.integer(sequence_number)) %>%
  arrange(seq)

ggplot(plot_data, aes(x = seq, y = as.numeric(home_win_percentage))) +
  geom_line(color = "#221A4D", linewidth = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey50") +
  labs(
    title = paste("Win probability --", game_id),
    x = "Play sequence",
    y = "Home team win probability"
  ) +
  theme_minimal()
```

### Odds and officials

```{r phase3-odds}
# WNBA odds -- populated when ESPN carries lines for this game
odds <- espn_wnba_event_odds(event_id = game_id)

# Officials list
officials <- espn_wnba_event_officials(event_id = game_id)
officials[, c("full_name", "position")]

# Broadcast info
broadcasts <- espn_wnba_event_broadcasts(event_id = game_id)
broadcasts[, c("market", "names")]
```

**WBB note.** ESPN does not carry odds for NCAA basketball games. Calling
`espn_wbb_event_odds()` returns an empty tibble -- this is expected and not
an error. The function exists for API surface symmetry so WBB and WNBA
workflows have an identical interface.

## Phase 4 walkthrough -- WNBA-specific

### Draft picks

```{r phase4-draft}
library(wehoop)

# 2025 WNBA draft
draft_2025 <- espn_wnba_draft(season = 2025)
head(draft_2025[, c("pick", "round", "team_name", "athlete_display_name")])
```

### Transactions

```{r phase4-transactions}
# Recent WNBA transactions (waiver claims, trades, signings)
txn <- espn_wnba_transactions(season = 2025, limit = 100)
head(txn[, c("date", "type", "athlete_display_name", "team_name")])
```

### Free agents

```{r phase4-fa}
# WNBA free agents -- returns empty tibble outside the free-agency window
fa <- espn_wnba_freeagents(season = 2025)
nrow(fa)
```

The `espn_wnba_freeagents()` call returns an empty tibble whenever the
league is not in its free-agency window. Check `nrow(fa) > 0` before
downstream processing.

## Phase 5 walkthrough -- league-wide catalogs

### Statistical leaders

```{r phase5-leaders}
library(wehoop)

# WNBA 2025 statistical leaders (points, rebounds, assists, ...)
wnba_ldrs <- espn_wnba_leaders(season = 2025, season_type = 2)
head(wnba_ldrs)

# WBB 2025 statistical leaders
wbb_ldrs <- espn_wbb_leaders(season = 2025, season_type = 2)
head(wbb_ldrs)
```

### Venues and coaches

```{r phase5-venues-coaches}
# All WNBA arenas (no season argument -- returns full historical list)
wnba_venues <- espn_wnba_venues()
head(wnba_venues[, c("full_name", "city", "state", "capacity")])

# All WBB coaches for 2025
wbb_coaches <- espn_wbb_coaches(season = 2025)
head(wbb_coaches[, c("full_name", "team_name", "experience")])
```

### Athletes index

The athletes index covers every player ESPN tracks for the league and season.
The WBB index is large (potentially 6,000+ athletes); use `limit` to cap the
initial pull or set `active = TRUE` to restrict to rostered players.

```{r phase5-athletes}
# WNBA 2025 active athletes
wnba_athletes <- espn_wnba_athletes_index(season = 2025, active = TRUE, limit = 5000)
nrow(wnba_athletes)
head(wnba_athletes[, c("display_name", "position_name", "team_name")])

# WBB -- cap at 50 rows for illustration; default limit = 25000
wbb_athletes <- espn_wbb_athletes_index(season = 2025, limit = 50)
head(wbb_athletes[, c("display_name", "position_name", "team_name")])
```

### Seasons catalog

```{r phase5-seasons}
# All WNBA seasons ESPN has on record
wnba_seasons <- espn_wnba_seasons()
head(wnba_seasons[, c("year", "start_date", "end_date")])

# Season metadata for 2025 (mostly $ref URLs to sub-resources)
wnba_s2025 <- espn_wnba_season_info(season = 2025)
head(wnba_s2025)

# Same for WBB
wbb_seasons <- espn_wbb_seasons()
wbb_s2025 <- espn_wbb_season_info(season = 2025)
```

### Phase 6 -- WNBA conferences

```{r phase6-conferences}
# WNBA conference list (parity with espn_wbb_conferences())
wnba_confs <- espn_wnba_conferences()
wnba_confs[, c("id", "name", "short_name")]
```

## Gaps and limitations

The following ESPN basketball endpoints are intentionally not wrapped by
wehoop:

| Endpoint | Reason |
|---|---|
| `/teams/{id}/depth-charts` | ESPN basketball does not carry depth charts |
| `cdn.espn.com/core` game summary | Fully redundant with existing `espn_*_game_all()` parsing |
| core-v2 `/events/{id}/plays` | Strict downgrade of `espn_*_pbp()`: requires per-play `$ref` resolution, returns less data |
| `espn_wnba_rankings` | WNBA has no poll or rankings surface; only WBB has `espn_wbb_rankings()` |
| WBB transfer portal | Not on ESPN basketball API |

Known data limitations by function:

- **`espn_wbb_injuries()`** and **`espn_wbb_team_injuries()`** -- ESPN rarely
  populates injury data for college basketball. An empty tibble is the common
  case, not an error.
- **`espn_wbb_event_odds()`** -- ESPN does not carry betting lines for NCAA
  games. Always returns an empty tibble.
- **`espn_wnba_freeagents()`** -- empty outside the free-agency window.
- **`espn_*_athlete_awards()`** -- sparse; most athletes have no
  ESPN-recorded awards.
- **`espn_*_athlete_eventlog()`** -- returns `$ref` URL columns, not parsed
  stats. Use `espn_*_athlete_gamelog()` for game statistics.
- **`espn_*_season_info()`** -- payload consists mostly of `$ref` URLs to
  sub-resources; use `espn_*_seasons()` for a more useful tabular view.
- **web-common-v3 endpoints** (`_athlete_overview`, `_athlete_stats`,
  `_athlete_gamelog`, `_athlete_splits`) -- less stable than site-v2; may
  return HTTP 404 for legacy seasons or less-prominent athletes.
- **`espn_*_athletes_index()`** for WBB -- the full index can be 10,000+
  rows for current seasons. The default `limit = 25000` paginates internally.
  Set `limit = 50` or `limit = 100` for quick exploratory pulls.
