---
title: "WNBA Stats API Parameter Descriptions"
description: "A searchable reference table of every query parameter used by the WNBA Stats API endpoints wrapped in <code>wehoop</code> -- parameter name, description, accepted values, and which functions accept it."
author: "Saiem Gilani <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/wehoop-data/blob/main/themes/wehoop_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>"
)
with_dt <- requireNamespace("DT")
```

```{r eval = with_dt}
DT::datatable(wehoop::parameter_descriptions, options = list(scrollX = TRUE, pageLength = 25), filter = "top", rownames = FALSE)
```

```{r eval = !with_dt}
knitr::kable(wehoop::parameter_descriptions)
```

# **Our Authors**
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>
-   [Geoffery Hutchinson](https://x.com/hutchngo)
<a href='https://x.com/hutchngo' target='blank'><img src='https://img.shields.io/twitter/follow/hutchngo?color=blue&amp;label=%40hutchngo&amp;logo=x&amp;style=for-the-badge' alt='@hutchngo'/></a>
<a href='https://github.com/hutchngo' target='blank'><img src='https://img.shields.io/github/followers/hutchngo?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@hutchngo'/></a>

## **Citation**

To cite the [**`wehoop`**](https://wehoop.sportsdataverse.org/) R package in publications, use:

BibTex Citation

```bibtex
@misc{wehoop,
  author = {Saiem Gilani, Geoffery Hutchinson},
  title = {wehoop: The SportsDataverse's R Package for Women's Basketball Data},
  url = {https://wehoop.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
