---
title: "Women's College Basketball Cookbook"
description: "A storytelling tour of women's college basketball analysis with <code>wehoop</code> -- recipes from a single game to a full bracket, plus NCAA NET rankings, and a running commentary on how to read the package's function names."
author: "Saiem Gilani, Geoffery Hutchinson <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://x.com/hutchngo' target='blank'><img src='https://img.shields.io/twitter/follow/hutchngo?color=blue&amp;label=%40hutchngo&amp;logo=x&amp;style=for-the-badge' alt='@hutchngo'/></a> <a href='https://github.com/hutchngo' target='blank'><img src='https://img.shields.io/github/followers/hutchngo?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@hutchngo'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/wehoop-data/blob/main/themes/wehoop_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = "#>", collapse = TRUE,
                      warning = FALSE, message = FALSE, error = TRUE)
```

## Before we cook

Women's college basketball has never been more watched, and the data has
never been better. `wehoop` covers it end to end -- 360-odd Division I
teams, a March bracket, and the NCAA's own NET rankings. The trick to
not drowning is the same as everywhere else in this family of packages:
**learn the grammar of the function names and you can guess your way to
almost anything.**

If you've read the WNBA cookbook, you already know most of the grammar.
This one swaps the league token from `wnba` to `wbb` and adds one new
prefix -- `ncaa_wbb_` for NCAA.com data. That's nearly the whole diff.

### The grammar, refreshed for college

A `wehoop` college function answers three questions, in order:

1. **Where's the data from?** -- the prefix.
   - `espn_` -- ESPN's APIs. The broad, stable backbone.
   - `ncaa_wbb_` -- NCAA.com. Home of the NET rankings and the official
     bracket machinery. (Here the prefix carries the league with it.)
   - `load_` -- pre-built bulk season files.
2. **Which league?** -- for ESPN it's spelled out: `espn_wbb_`. The
   `wbb` token is "women's basketball."
3. **What do you want?** -- the rest of the name, general to specific.

So "a team's schedule from ESPN" is `espn_` + `wbb_` + `team_schedule`
-> `espn_wbb_team_schedule()`. "NCAA NET rankings" is `ncaa_wbb_` +
`NET_rankings` -> `ncaa_wbb_NET_rankings()`. You'll be right far more
often than wrong, and being wrong just means opening the reference
index.

```{r}
library(wehoop)
library(dplyr)
```

## Recipe 1: A powerhouse program's season

**The story.** It's November and you want to set the table for UConn's
season -- schedule, roster, where they sit.

Type `espn_wbb_team` and the grammar lays out the menu:

```{r}
team_id <- 2509            # UConn
season  <- most_recent_wbb_season()

espn_wbb_team(team_id = team_id)                              # identity + record
espn_wbb_team_schedule(team_id = team_id, season = season)    # the slate
espn_wbb_team_roster(team_id = team_id, season = season)      # the players
espn_wbb_team_season_statistics(team_id = team_id, season = season)
```

Margin note: `most_recent_wbb_season()` is worth knowing. Every league
in `wehoop` has a `most_recent_<league>_season()` companion, and the
newer `espn_wbb_*` functions use it as their default `season`. Most of
the time you can simply *omit* `season` -- the function already knows
what "now" means.

## Recipe 2: The star guard

**The story.** Women's college basketball has become a star-driven
sport, and you want a marquee guard's numbers.

College rosters turn over fast -- this year's star is next year's pro --
so the reliable move is to pull the current roster first and lift an
`athlete_id` straight off it, then walk the `athlete_` family:

```{r}
roster     <- espn_wbb_team_roster(team_id = 2509, season = most_recent_wbb_season())
athlete_id <- roster$athlete_id[1]   # whoever is first on the roster

espn_wbb_player_info(athlete_id = athlete_id)            # bio
espn_wbb_player_gamelog(athlete_id = athlete_id, season = most_recent_wbb_season())
espn_wbb_player_career_stats(athlete_id = athlete_id)    # career rollup, long format
```

That "pull the index, lift an ID, drill down" move is the single most
useful habit in the whole package. `athlete_gamelog` in particular only
has data for players *currently* on a roster -- last year's star has
moved on -- so feeding it an ID you just lifted off `team_roster` is how
you guarantee it returns something.

Margin note: `athlete_info`, `athlete_gamelog`, `athlete_career_stats`
all share the `athlete_` stem. A repeated stem is a *namespace* -- a
promise that everything under it is about the same subject. Browsing by
stem (`espn_wbb_athlete` + Tab) is often faster than searching.

## Recipe 3: The NET rankings

**The story.** Box scores and the eye test only get you so far in
seeding arguments. The NCAA's NET ranking is the number that actually
drives the bracket, and you want it.

This is where the new prefix earns its keep. NCAA.com data is
`ncaa_wbb_`.

```{r}
# ncaa_wbb / NET / rankings. The name is the thing.
net <- ncaa_wbb_NET_rankings()

net |>
  select(rank, team, conference, net_wins, net_losses) |>
  head(25)
```

The grammar lesson: a prefix is a compressed paragraph. The moment you
see `ncaa_wbb_`, you know the data is women's-college-only and comes
from the NCAA's own systems rather than ESPN's. Learn what each prefix
promises and a lot of the documentation becomes redundant.

## Recipe 4: Game flow

**The story.** A December non-conference game went down to the wire.
You want the play-by-play and a sense of the swing.

Same three zoom levels as every other cookbook -- the grammar is
league-agnostic:

```{r}
game_id <- 401276115

espn_wbb_game_all(game_id = game_id)      # everything, as a named list
espn_wbb_team_box(game_id = game_id)      # team box
espn_wbb_player_box(game_id = game_id)    # player box
espn_wbb_pbp(game_id = game_id)           # play-by-play

# ESPN's win-probability-per-play for the same game.
espn_wbb_game_probabilities(event_id = game_id)
```

If you read the WNBA cookbook, you've now seen `espn_*_game_all`,
`espn_*_team_box`, `espn_*_player_box`, `espn_*_pbp` in two leagues with
*only the league token changing*. That's the payoff of a regular
grammar: a recipe learned once transfers by editing a single word.

## Recipe 5: Per-player, per-game box scores

**The story.** You want one player's line from one game, tidy and long
-- not the whole `player_box`.

```{r}
# event + team + athlete -- three IDs, because you're naming one cell
# of one game. More identifiers = finer grain. Always.
espn_wbb_game_player_box(
  event_id   = 401276115,
  team_id    = 52,
  athlete_id = 4398583
)
```

## Recipe 6: The conference race

**The story.** It's February and your conference is a dogfight. You want
the standings and the structure beneath them.

College conferences are a *hierarchy* -- divisions, conferences, the
whole D-I grouping -- and ESPN models that with the `season_group`
family.

```{r}
season <- most_recent_wbb_season()

# A grammar gotcha worth flagging: espn_wbb_standings() takes `year`,
# not `season`. Most functions say `season`, a few older ones say
# `year` -- when an argument name surprises you, ?function is the
# two-second fix.
espn_wbb_standings(year = season)

# The group tree: groups -> one group -> its teams.
espn_wbb_season_groups(season = season, season_type = 2)
espn_wbb_season_group(group_id = 50, season = season, season_type = 2)
espn_wbb_season_group_teams(group_id = 50, season = season, season_type = 2)
```

Margin note: `season_groups` (plural) is the index; `season_group`
(singular) is one item. That singular/plural pair runs through the whole
package -- `franchises`/`franchise`, `tournaments`/`tournament`,
`positions`/`position`. Plural = "give me the list," singular = "give me
this one." It's the most reliable two-word rule in the grammar.

## Recipe 7: March

**The story.** The bracket. You want the tournament.

```{r}
season <- most_recent_wbb_season()

espn_wbb_tournaments()                                  # which tournaments exist
espn_wbb_tournament(tournament_id = 3)                  # one tournament
espn_wbb_tournament_seasons(tournament_id = 3)          # its yearly editions
espn_wbb_tournament_season(tournament_id = 3, season = season)
```

Same singular/plural rule as Recipe 6, same general-to-specific march
from `tournaments` down to one `tournament_season`. By now you should be
able to *predict* this family before reading it -- which is the whole
point of the cookbook.

## Recipe 8: A whole season, in bulk

**The story.** You're done with single games. You want every women's
college game of a season to model on.

Looping `espn_wbb_pbp()` over thousands of games works but is slow and
rude to ESPN. The `load_` prefix exists for exactly this:

```{r eval = FALSE}
# load / wbb / pbp -- one call, one season.
pbp <- load_wbb_pbp(seasons = 2024)

load_wbb_player_box(seasons = 2024)
load_wbb_team_box(seasons = 2024)
load_wbb_schedule(seasons = 2024)
load_wbb_shots(seasons = 2024)         # shot locations, ready to chart
```

And straight into a database, with the `(seasons, dbConnection,
tablename)` plumbing every `load_` function shares:

```{r eval = FALSE}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "wbb.sqlite")
load_wbb_pbp(seasons = 2020:2024, dbConnection = con, tablename = "wbb_pbp")
dbDisconnect(con)
```

Margin note: consistency is a feature. Every `load_` function shares
the same signature, so learning one teaches you all of them -- guess
the arguments the same way you guess the names.

## Working through a proxy

Campus networks love a proxy. `wehoop` handles them in three layers --
reach for the least invasive one that works.

**Layer 1 -- set it once for the session.** Covers everything: ESPN,
NCAA, loaders.

```{r eval = FALSE}
options(wehoop.proxy = "http://proxy.university.edu:8080")

# Authenticated proxy? Pass a list instead of a string.
options(wehoop.proxy = list(
  url      = "http://proxy.university.edu",
  port     = 8080,
  username = "netid",
  password = Sys.getenv("PROXY_PASS")
))

# Then just work -- every call inherits it.
espn_wbb_team_roster(team_id = 2509, season = most_recent_wbb_season())
ncaa_wbb_NET_rankings()
```

**Layer 2 -- environment variables.** For shared scripts and CI, where
the proxy belongs to the machine and not the analysis, set the standard
variables and let libcurl pick them up:

```{r eval = FALSE}
Sys.setenv(
  http_proxy  = "http://proxy.university.edu:8080",
  https_proxy = "http://proxy.university.edu:8080",
  no_proxy    = "localhost,127.0.0.1"
)
```

A grammar-flavored note on what *doesn't* take a per-call `proxy =`
argument: the `espn_*`, `ncaa_wbb_*`, and `load_*` functions call the
HTTP layer directly, so they read the proxy from the session option or
the environment -- not from an argument. (In the WNBA cookbook, the
`wnba_*` Stats API functions *do* accept per-call `proxy =`, because
they thread `...` through. Women's college basketball has no equivalent
Stats-API surface, so for `wbb` you'll use Layer 1 or Layer 2 every
time.) Once again, the prefix tells you the capability before you go
looking.

## Where to go next

The women's college game has more teams, a deeper conference hierarchy,
and the `ncaa_wbb_` prefix -- but the grammar is identical to every
other league in this family. You decided where the data lived, you
named the league, you named the thing from general to specific, and you
let the singular/plural rule and the shared stems carry you the rest of
the way.

`wehoop` is the women's-game sibling of `hoopR`, and the two share this
grammar bone for bone. `espn_wbb_team_roster` has a mirror image in
`hoopR`'s `espn_mbb_team_roster`; a recipe you learn in one package
crosses straight into the other. Four leagues, one grammar -- and now
it's yours.

# **Our Authors**
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>
-   [Geoffery Hutchinson](https://x.com/hutchngo)
<a href='https://x.com/hutchngo' target='blank'><img src='https://img.shields.io/twitter/follow/hutchngo?color=blue&amp;label=%40hutchngo&amp;logo=x&amp;style=for-the-badge' alt='@hutchngo'/></a>
<a href='https://github.com/hutchngo' target='blank'><img src='https://img.shields.io/github/followers/hutchngo?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@hutchngo'/></a>

## **Citation**

To cite the [**`wehoop`**](https://wehoop.sportsdataverse.org/) R package in publications, use:

BibTeX Citation

```bibtex
@misc{wehoop,
  author = {Saiem Gilani and Geoffery Hutchinson},
  title = {wehoop: The SportsDataverse},
  url = {https://wehoop.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
