---
title: "WNBA Cookbook"
description: "A storytelling tour of WNBA analysis with <code>wehoop</code> -- recipes from a single box score to a full season, with a running commentary on how the package's function names are built so you can guess the next one."
author: "Saiem Gilani, Geoffery Hutchinson <br><a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a> <a href='https://x.com/hutchngo' target='blank'><img src='https://img.shields.io/twitter/follow/hutchngo?color=blue&amp;label=%40hutchngo&amp;logo=x&amp;style=for-the-badge' alt='@hutchngo'/></a> <a href='https://github.com/hutchngo' target='blank'><img src='https://img.shields.io/github/followers/hutchngo?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@hutchngo'/></a>"
opengraph:
  image:
    src: "https://github.com/sportsdataverse/wehoop-data/blob/main/themes/wehoop_gh.png?raw=true"
  twitter:
    card: summary_large_image
    creator: "@saiemgilani"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = "#>", collapse = TRUE,
                      warning = FALSE, message = FALSE, error = TRUE)
```

## Before we cook

Every recipe here is a small story: a real question about the WNBA, and
the shortest honest path from that question to a tidy data frame. Most
of the code blocks run live when this page builds, so the tables you see
are real data freshly pulled from ESPN and WNBA.com. A handful -- bulk
season downloads, the database recipe, and the proxy examples -- are
shown but not run, since they either take too long or would reconfigure
your session; those are flagged where they appear.

The one idea to carry through all of it: **`wehoop` function names are a
grammar, not a list to memorize.** Learn the grammar and you can *guess*
the function you need. That guessing skill is what this cookbook is
really teaching; the basketball is the delivery vehicle.

### The grammar of a `wehoop` function name

Almost every function answers three questions, in order:

1. **Where does the data come from?** -- the prefix.
   - `espn_` -- ESPN's APIs. The broadest, most stable coverage.
   - `wnba_` -- the official WNBA Stats API (stats.wnba.com). Deepest
     box scores, tracking data, lineup data.
   - `load_` -- pre-built bulk data. A whole season in one call.
2. **Which league?** -- for ESPN functions the league is spelled out:
   `espn_wnba_`. For the WNBA Stats API the `wnba_` prefix *is* the
   league.
3. **What do you want?** -- the rest of the name, read left to right,
   general to specific: `team_season_roster`, `athlete_career_stats`,
   `event_competitor_linescores`.

So "a team's roster for a season, from ESPN" assembles in your head as
`espn_` + `wnba_` + `team` + `_roster` -> `espn_wnba_team_roster()`. "The
league leaders from WNBA.com" is `wnba_` + `leagueleaders` ->
`wnba_leagueleaders()`. Hundreds of functions, one grammar. Memorize the
grammar, not the list.

We'll point at the grammar in the margins of every recipe.

```{r}
library(wehoop)
library(dplyr)
```

## Recipe 1: One game, three levels of detail

**The story.** You caught the end of a game on a stream and want to
actually study it -- not the recap, the game.

ESPN serves a game at three zoom levels, and the function name tells you
which zoom you asked for. Start wide, push in.

```{r}
game_id <- 401736171

# Widest: the entire game summary as a named list of tibbles.
# Read the name: espn / wnba / "game" / "all".
full_game <- espn_wnba_game_all(game_id = game_id)
names(full_game)

# Medium: the box score. team_box and player_box are siblings -- same
# noun (box), different grain.
espn_wnba_team_box(game_id = game_id)
espn_wnba_player_box(game_id = game_id)

# Tightest: play-by-play, one row per possession.
espn_wnba_pbp(game_id = game_id)
```

Margin note: `team_box` and `player_box` differ by exactly one word, and
that word is the *grain* of the table. Two functions that differ by a
single noun -- that noun is telling you "one row per ___."

## Recipe 2: A franchise cornerstone's whole career

**The story.** You want A'ja Wilson's career arc -- all of it, tidy
enough to plot -- not just this season.

The word you want is **career**.

```{r}
# espn / wnba / athlete / career / stats. The name is the sentence.
aja <- espn_wnba_player_career_stats(athlete_id = 3149391)

# It returns long -- one row per (stat_type x category x stat) --
# exactly the shape a faceted ggplot wants.
aja |>
  filter(category_name == "offensive", stat_name == "avgPoints") |>
  select(stat_type_id, value)
```

Margin note: `athlete_career_stats()` defaults to pulling *both* regular
season and postseason and stacking them, keyed by `stat_type_id`. Many
of the newer `espn_wnba_*` functions behave this way -- when "regular +
playoffs" both make sense, you get both and filter. Want just one? Pass
`season_type`.

The season-by-season index is one word away -- swap `career_stats` for
`seasons`:

```{r}
espn_wnba_player_seasons(athlete_id = 3149391)
```

## Recipe 3: Game flow and the fourth-quarter swing

**The story.** That game from Recipe 1 turned on a late run. You want to
*see* it -- a win-probability line.

Two engines, and the prefix tells you which:

```{r}
# Route A -- ESPN's per-play win probability.
espn_wnba_game_probabilities(event_id = 401736171)

# Route B -- the V3 play-by-play from the WNBA Stats API itself. The
# prefix flips to `wnba_` -> different data source, different first
# token. It keys off WNBA.com game ids, which look nothing like ESPN's,
# so lift a real one off the schedule first.
sched <- wnba_schedule()
wnba_playbyplayv3(game_id = sched$game_id[1])
```

The grammar doing real work: when two functions look like near-synonyms,
the *prefix* is the tiebreaker. `espn_wnba_*` is ESPN's take; `wnba_*`
is WNBA.com's. Choose by which columns you want, not by which name you
remembered.

## Recipe 4: One player's line from one game

**The story.** You want a single athlete's box score from a single game,
long and tidy -- not the whole `player_box`.

```{r}
# espn / wnba / event / player_box. You pass the (event, team, athlete)
# trio because that triple is what uniquely names "this player in this
# game."
espn_wnba_game_player_box(
  event_id   = 401736171,
  team_id    = 17,
  athlete_id = 2590093
)
```

Margin note on *why three IDs*: a function's argument list is a hint
about its grain. One ID (`game_id`) -> a whole game. Three IDs
(`event` + `team` + `athlete`) -> one cell of it. More identifiers, finer
grain -- the arguments and the grain always move together.

The event-detail siblings all share the `espn_wnba_event_` stem:

```{r}
espn_wnba_game_team_linescores(event_id = 401736171, team_id = 17)
espn_wnba_game_officials(event_id = 401736171)
espn_wnba_game_play(event_id = 401736171, play_id = 4017361714)
```

## Recipe 5: A team's season at a glance

**The story.** You're previewing a team -- record, schedule, roster, and
how they really stack up.

Type `espn_wnba_team` and let the grammar lay out the menu:

```{r}
team_id <- 17    # Las Vegas Aces
season  <- most_recent_wnba_season()

espn_wnba_team_season_profile(team_id = team_id, season = season)
espn_wnba_team_schedule(team_id = team_id, season = season)
espn_wnba_team_roster(team_id = team_id, season = season)

# The full team-season stat sheet (with league rank per stat) vs. just
# the W-L record.
espn_wnba_team_season_statistics(team_id = team_id, season = season)
espn_wnba_team_record(team_id = team_id, season = season)
```

Notice the family: `team_season_profile`, `team_season_roster`,
`team_season_statistics`. The shared middle (`team_season`) is the
subject; the last word is the lens. Find one member, you've found the
family.

## Recipe 6: The league-leaders board

**The story.** You want the scoring, rebounding, and assist leaders --
the bones of every MVP argument.

```{r}
# WNBA.com's leaders. Prefix `wnba_` -> the official Stats API.
wnba_leagueleaders(season = "2024")

# ESPN's version of the same idea. Same question, different kitchen --
# the prefix tells you which before you read another word.
espn_wnba_leaders(season = most_recent_wnba_season())
```

Leaders live under *both* prefixes because both APIs publish them. That
is normal and intentional -- `wehoop` mirrors the data sources honestly
rather than hiding one. You pick based on which columns you want.

## Recipe 7: Draft night

**The story.** You want to follow a draft class -- order, picks,
players.

The draft has its own little family, sorted general to specific exactly
as the grammar promises:

```{r eval = FALSE}
season <- most_recent_wnba_season()

espn_wnba_season_draft(season = season)        # the draft, top level
espn_wnba_draft_rounds(season = season)        # by round
espn_wnba_draft_athletes(season = season)      # every drafted player
espn_wnba_draft_pick(season = season, round = 1, pick = 1)

# Or the pre-built bulk version -- note the `load_` prefix.
load_wnba_draft(seasons = 2020:2024)
```

Read it top to bottom: each step adds a word and narrows the grain.
That's the grammar's "general to specific, left to right" rule made
visible.

## Recipe 8: A whole season at once

**The story.** You don't want one game. You want every game -- a full
season to model on.

Looping `espn_wnba_pbp()` over hundreds of `game_id`s works but is slow
and rude to ESPN. The `load_` prefix exists for exactly this: pre-built,
pre-cleaned season files.

```{r eval = FALSE}
# load / wnba / pbp -- one call, one season.
pbp_2024 <- load_wnba_pbp(seasons = 2024)

load_wnba_player_box(seasons = 2024)
load_wnba_team_box(seasons = 2024)
load_wnba_schedule(seasons = 2024)
load_wnba_shots(seasons = 2024)        # shot locations, ready to chart
```

There's also a deeper, WNBA-Stats-API-sourced bulk family under the
`load_wnba_stats_` stem -- richer play-by-play with on-court lineups,
possession assignment, and a garbage-time flag:

```{r eval = FALSE}
load_wnba_stats_pbp(seasons = 2024)
load_wnba_stats_player_game_logs(seasons = 2024)
load_wnba_stats_leaguedash(seasons = 2024, table = "lineups_master")
```

Every `load_wnba_stats_` family is published back to **1997**, with two
documented exceptions: officials start in 2004, and the coaches tag has no
2019 season. Pass `seasons = TRUE` to any of them for the full history.

The grammar lesson: `load_wnba_` and `load_wnba_stats_` are two bulk
families, and the extra `stats_` token tells you the *source* -- the
plain family is ESPN-shaped, the `stats_` family is WNBA.com-shaped. An
extra token in the middle of a name is never noise; it's always
narrowing something.

## Recipe 9: Pour a season straight into a database

**The story.** A season of play-by-play is big. You'd rather query it
from disk than hold it in memory.

Every `load_` function shares the same optional plumbing -- learn it
once, you've learned it everywhere:

```{r eval = FALSE}
library(DBI)
con <- dbConnect(RSQLite::SQLite(), "wnba.sqlite")

load_wnba_pbp(
  seasons      = 2020:2024,
  dbConnection = con,
  tablename    = "wnba_pbp"
)

dbDisconnect(con)
```

Margin note: consistency is a feature. Because every `load_` function
shares the `(seasons, ..., dbConnection, tablename)` signature, learning
one teaches you all of them -- guess the arguments the same way you
guess the names.

## Working through a proxy

On a corporate or campus network your traffic may need a proxy.
`wehoop` understands proxies in three layers -- use the least invasive
one that works.

**Layer 1 -- set it once for the session.** This covers everything:
ESPN, WNBA Stats, loaders.

```{r eval = FALSE}
# Plain proxy.
options(wehoop.proxy = "http://proxy.company.com:8080")

# Authenticated proxy -- pass a list and it's spread into the request.
options(wehoop.proxy = list(
  url      = "http://proxy.company.com",
  port     = 8080,
  username = "saiem",
  password = Sys.getenv("PROXY_PASS")
))

# From here on, nothing changes -- just call functions normally.
espn_wnba_team_roster(team_id = 17, season = most_recent_wnba_season())
wnba_leagueleaders(season = "2024")
```

**Layer 2 -- one call, one proxy.** Every WNBA Stats API function
(`wnba_*`) threads `...` down to the HTTP layer, so you can override the
proxy for a single call:

```{r}
wnba_playbyplayv3(
  game_id = "1022400001",
  proxy   = "http://other-proxy.company.com:3128"
)
```

A grammar-flavored caveat: per-call `proxy =` works for the `wnba_*`
Stats API functions because they pass `...` through. The `espn_*` and
`load_*` functions call the HTTP layer directly and don't take a
per-call proxy -- for those, use the session option in Layer 1. The
prefix tells you the capability: `wnba_` functions are the ones wired
for per-call overrides.

**Layer 3 -- let the environment decide.** If you'd rather keep proxies
out of your R code entirely, `wehoop` honors the standard environment
variables:

```{r eval = FALSE}
Sys.setenv(
  http_proxy  = "http://proxy.company.com:8080",
  https_proxy = "http://proxy.company.com:8080",
  no_proxy    = "localhost,127.0.0.1"
)
```

libcurl reads these automatically whenever you haven't set an explicit
proxy. It's the right layer for shared scripts and CI, where the proxy
is a property of the machine, not the analysis.

## Where to go next

You have the grammar now. Every recipe above was the same move: decide
where the data lives (the prefix), name the league, then name the thing
general to specific. The WNBA surface in `wehoop` is large but
*regular* -- and regular things are guessable.

The women's college game runs on the same package with the same
grammar -- swap `wnba` for `wbb`, meet the NCAA-sourced functions, and
keep cooking. That's the WBB cookbook.

# **Our Authors**
-   [Saiem Gilani](https://x.com/saiemgilani)
<a href='https://x.com/saiemgilani' target='blank'><img src='https://img.shields.io/twitter/follow/saiemgilani?color=blue&amp;label=%40saiemgilani&amp;logo=x&amp;style=for-the-badge' alt='@saiemgilani'/></a>
<a href='https://github.com/saiemgilani' target='blank'><img src='https://img.shields.io/github/followers/saiemgilani?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@saiemgilani'/></a>
-   [Geoffery Hutchinson](https://x.com/hutchngo)
<a href='https://x.com/hutchngo' target='blank'><img src='https://img.shields.io/twitter/follow/hutchngo?color=blue&amp;label=%40hutchngo&amp;logo=x&amp;style=for-the-badge' alt='@hutchngo'/></a>
<a href='https://github.com/hutchngo' target='blank'><img src='https://img.shields.io/github/followers/hutchngo?color=eee&amp;logo=Github&amp;style=for-the-badge' alt='@hutchngo'/></a>

## **Citation**

To cite the [**`wehoop`**](https://wehoop.sportsdataverse.org/) R package in publications, use:

BibTeX Citation

```bibtex
@misc{wehoop,
  author = {Saiem Gilani and Geoffery Hutchinson},
  title = {wehoop: The SportsDataverse},
  url = {https://wehoop.sportsdataverse.org/},
  year = {2026}
}
```

## **Related SportsDataverse packages**

-   [**cfbfastR**](https://cfbfastR.sportsdataverse.org/) - college football
-   [**hoopR**](https://hoopR.sportsdataverse.org/) - men's basketball
-   [**wehoop**](https://wehoop.sportsdataverse.org/) - women's basketball
-   [**baseballr**](https://baseballr.sportsdataverse.org/) - baseball
-   [**fastRhockey**](https://fastRhockey.sportsdataverse.org/) - hockey
-   [**oddsapiR**](https://oddsapiR.sportsdataverse.org/) - betting odds
-   [**sportyR**](https://sportyR.sportsdataverse.org/) - playing surfaces
-   [**sportsdataverse-py**](https://py.sportsdataverse.org/) - the Python package
-   [**sportsdataverse-R**](https://r.sportsdataverse.org/) - the R meta-package
